---
title: "Project"
author: "Thayumanav Jayadevan"
date: "December 3, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import libraries
```{r}
library(quantmod)
library(PerformanceAnalytics)
library(fPortfolio)
library(quadprog)
library(tseries)
```

### Get data from Yahoo Finance
```{r}
data.INTC = getSymbols("INTC", from = "2011-11-30", to = "2018-11-30", auto.assign = F)
data.MCD = getSymbols("MCD", from = "2011-11-30", to = "2018-11-30", auto.assign = F)
data.TM = getSymbols("TM", from = "2011-11-30", to = "2018-11-30", auto.assign = F)
data.NKE = getSymbols("NKE", from = "2011-11-30", to = "2018-11-30", auto.assign = F)
data.WMT = getSymbols("WMT", from = "2011-11-30", to = "2018-11-30", auto.assign = F)

data.GSPC = getSymbols("^GSPC", from = "2011-11-30", to = "2018-11-30", auto.assign = F)
```

# Returns

### Calculate returns of stocks
```{r}
options(scipen = 4)

data.INTC$INTC.ret = Return.calculate(data.INTC$INTC.Adjusted)
data.MCD$MCD.ret = Return.calculate(data.MCD$MCD.Adjusted)
data.TM$TM.ret = Return.calculate(data.TM$TM.Adjusted)
data.NKE$NKE.ret = Return.calculate(data.NKE$NKE.Adjusted)
data.WMT$WMT.ret = Return.calculate(data.WMT$WMT.Adjusted)

data.GSPC$GSPC.ret = Return.calculate(data.GSPC$GSPC.Adjusted)
```

### Construct portfolio
```{r}
multi = cbind(data.INTC$INTC.ret, data.MCD$MCD.ret, data.TM$TM.ret, data.NKE$NKE.ret, 
              data.WMT$WMT.ret, data.GSPC$GSPC.ret)
multi = multi[-1, ]
```

### Portfolio return equal weighted, rebalanced on quarters
```{r}
port.ret = Return.portfolio(multi, rebalance_on = "quarters")
```

```{r}
Return.cumulative(port.ret)
```

# Risk
### Create a matrix and transposed matrix of equal weights for 6 stocks
```{r}
weights = rep(1/6, 6)
mat.weight = matrix(weights, 1)
tmat.weight = t(mat.weight)
```

### Convert portfolio returns to matrix form
```{r}
mat.return = matrix(multi, nrow = nrow(multi))
```

### Create Covariance matrix
```{r}
VCOV = cov(mat.return) * 252
```

### Calculate portfolio risk
```{r}
port.var = mat.weight %*% VCOV %*% tmat.weight
port.sd = sqrt(port.var)
```

## Value at risk
## Gaussian VaR
### Calculate mean return and risk
```{r}
port.mean = mean(port.ret)
port.risk = sd(port.ret)
```

### Calculate current portfolio value
Assume we had invested is $50,000
```{r}
ret.cum = Return.cumulative(port.ret)
current_val = 50000 * (1 + ret.cum)
```

### Gaussian VaR at 1% and 5%
```{r}
VaR01.Gaussian = -(port.mean+port.risk*qnorm(0.01))*current_val
VaR01.Gaussian
VaR05.Gaussian = -(port.mean+port.risk*qnorm(0.05))*current_val
VaR05.Gaussian
```

## Historical VaR
```{r}
ret.cum = data.frame(lapply(multi, FUN = Return.cumulative))
```

```{r}
ret.val = (1 + ret.cum) * (50000 / 6)
```

```{r}
sim.portPnL = ret.val$INTC.ret * multi$INTC.ret + ret.val$MCD.ret * multi$MCD.ret + 
              ret.val$TM.ret * multi$TM.ret + ret.val$NKE.ret * multi$NKE.ret + 
              ret.val$WMT.ret * multi$WMT.ret + ret.val$GSPC.ret * multi$GSPC.ret
names(sim.portPnL) <- "Port.PnL"
```

### Historical VaR at 1% and 5%
```{r}
VaR01.Historical = quantile(-sim.portPnL$Port.PnL,0.99)
VaR01.Historical
VaR05.Historical = quantile(-sim.portPnL$Port.PnL,0.95)
VaR05.Historical
```

### Plot VaR in relation to P&L density
```{r}
ret.d = density(sim.portPnL)
```

```{r}
plot(ret.d, xlab = "Profit & Loss", ylab = "", yaxt = "n", 
main = "Density of Simulated Portfolio P&L Over Three Years 
     And 1% and 5% 1-Day Historical Value-at-Risk (VaR)")
abline(v = -quantile(-sim.portPnL, 0.99), col = "gray", lty = 1)
abline(v = -quantile(-sim.portPnL, 0.95),col = "black",lty = 2)
x = seq(min(sim.portPnL), max(sim.portPnL), length = 1000)
y = dnorm(x, mean = mean(sim.portPnL), sd = sd(sim.portPnL))
lines(x, y, type = "l", col = "black", lwd = 1, lty = 3)
legend("topright", c("Simulated P&L Distribution", "Normal Distribution", "1% 1-Day VaR","5% 1-Day VaR"), 
       col = c("black","black","gray","black"), lty = c(1,3,1,2))
```

# Expected Shortfall
### Gaussian ES at 1% and 5%
```{r}
ES01.Gaussian = current_val * (port.mean + port.risk * (dnorm(qnorm(.01)) / .01))
ES01.Gaussian
ES05.Gaussian = current_val * (port.mean + port.risk * (dnorm(qnorm(.05)) / .05))
ES05.Gaussian
```

### Historical ES at 1% and 5%
```{r}
ES.PnL = sim.portPnL$Port.PnL
ES.PnL$dummy01 = ifelse(ES.PnL$Port.PnL < (-VaR01.Historical) , 1, 0)
ES.PnL$dummy05 = ifelse(ES.PnL$Port.PnL < (-VaR05.Historical) ,1 , 0)
```

```{r}
shortfall01 = subset(ES.PnL, ES.PnL$dummy01==1)
shortfall05 = subset(ES.PnL, ES.PnL$dummy05==1)
ES01.Historical =  -mean(shortfall01$Port.PnL)
ES01.Historical
ES05.Historical =  -mean(shortfall05$Port.PnL)
ES05.Historical
```

# Risk-Adjusted Portfolio Performance Measures
```{r}
Rf = .0007
# minimum acceptable return
mar = 0.015
```

```{r}
annual.port.ret = mean(port.ret$portfolio.returns) * 252
annual.port.sd = sd(port.ret$portfolio.returns)*sqrt(252)
```

### Sharpe ratio
```{r}
Sharpe.port = (annual.port.ret - Rf) / annual.port.sd
Sharpe.port
```

### Roy's safety first ratio
```{r}
Roy.SF.port = (annual.port.ret - mar) / annual.port.sd
Roy.SF.port
```

### Treynor ratio
```{r}
port.reg = lm(port.ret$portfolio.returns ~ multi$GSPC.ret, data = port.ret)
port.beta = port.reg$coefficients[2]
```

```{r}
Treynor.port = (annual.port.ret - Rf ) / port.beta
names(Treynor.port) = "Treynor ratio"
Treynor.port
```

### Sortino ratio
```{r}
period.mar = mar / 252
downside.port = port.ret
downside.port$dummy = ifelse(port.ret$portfolio.returns < period.mar, 1, 0)
```

```{r}
downside.port = subset(downside.port, downside.port$dummy == 1)
dd.port = sd(downside.port$portfolio.returns) * sqrt(252)
```

```{r}
Sortino.port = (annual.port.ret - mar) / dd.port
Sortino.port
```

### Information ratio
```{r}
Act.Ret.port = port.ret$portfolio.returns - multi$GSPC.ret
alpha.port = mean(Act.Ret.port) * 252
```

```{r}
tracking.error.port = sd(Act.Ret.port) * sqrt(252)
```

```{r}
IR.port = alpha.port / tracking.error.port
IR.port
```
